---
title: "ENM RSD"
author: "Laura"
date: "2023-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
 library(rJava)
options(java.parameters = "-Xmx8g")
library(SDMtune)
library(raster)
library(dismo)
library(sp)
library(ENMTools)
library(ntbox)
library(spam)
library(dotCall64)
library(grid)
library(fields)
library(viridis)
library(viridisLite)
library(readxl)
library(spThin)
library(leaflet)
library(plotROC)
```

Remove spatial correlation of plant disease presence points 
```{r}
RSD_ = read_xlsx("E:/Laura/TESIS/DATOS/RSD.xlsx")

clean = thin(
  loc.data = RSD_,
  lat.col = "y",
  long.col = "x",
  spec.col = "Especie",
  thin.par = 5,
  reps = 1,
  write.files = T,
  out.dir = './',
  out.base = "",
  locs.thinned.list.return = T,
  verbose = TRUE,
  write.log.file = F
)

RSD = clean[[1]]
```

Plot presence points
```{r}
leaflet(RSD)%>% 
  addTiles() %>%
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude)
```

ntbox allows to select which environmental layers are related with the presence points
```{r}
run_ntbox()
```

load and plot environmental layers 
```{r}
bc = list.files("E:/Laura/TESIS/RSD/CAPAS/ACTUAL", pattern = ".tif", full.names = T, all.files = T)
env = stack(bc)
env = check.env(env, verbose = TRUE)
```

```{r}
env1 = check.env(env, verbose = T)
names(env1)
```

change environmental layers names
```{r}
names(env1) = c("Annual Mean Temperature", "Annual Precipitation",
               "Precipitation of Driest Month", 
               "Precipitation Seasonality", "Precipitation of Driest Quarter",
               "Mean Diurnal Range",  "Isothermality", "Temperature Seasonality",
               "Max Temperature of Warmest Month", "Min Temperature of Coldest Month")
plot(env1)
```


```{r}
raster.cor.matrix(env, method = "pearson")
raster.cor.plot(env, method = "pearson")
```

Create the species object
```{r}
set.seed(2023)

thin = read.csv("E:/Laura/TESIS/RSD/_thin1.csv")

spRSD <- prepareSWD(
  species = "RSD", 
  p = thin, 
  a = dismo::randomPoints(env, 87), 
  env = env
)
```

Model train
```{r}
model <- train(method = "Maxent",
               data = spRSD,
               verbose = T)
```

```{r}
set.seed(2023)
datasets <-
  trainValTest(
    spRSD, 
    test = 0.5,
    only_presence = T,
  )

train <- datasets[[1]]
test <- datasets[[2]]
```

Model optimization
```{r}
h = list(
    reg = seq(0.01, 2, 0.01),
    fc = c("l","q","p","t","lq","h","lp","lt","lh","qp","qt","qh","pt","ph","th","lqp","lqt","lqh","lpt","lph","lth","qpt","qph","qth","pth","lqpt","lqph","lqth","lpth","qpth","lqpth"
    ),
    iter = seq(100, 500, 100)
  )
```

```{r}
pars = optimizeModel(model, 
              hypers = h, 
              metric = "aicc",
              test = test, 
              pop = 5, 
              gen = 5,
              env = env)
```

Select best model
```{r}
pars@results
#pars@models
pars@models[[1]]
```

```{r}
modelRSD <- train(method = "Maxent",
               data = train,
               verbose = T,
               fc = "qt", 
               reg = 0.56,
               iter = 400)
modelRSD
```


Best model metrics
```{r}
auc(modelRSD)
auc(modelRSD, test = test)
tss(modelRSD)
tss(modelRSD, test = test)
aicc(modelRSD, env)
```

Importance feature 
```{r}
VAR_I = varImp(modelRSD)
VAR_I
plotVarImp(VAR_I) 
```

Load environmental layers for projected areas
```{r}
env2 = list.files("E:/Laura/TESIS/RSD/CAPAS/ACTUAL1", pattern = ".tif", full.names = T, all.files = T)
env2 = stack(env2) 
env2 = check.env(env2, verbose = T)

bc126 = list.files("E:/Laura/TESIS/RSD/CAPAS/126", pattern = ".tif", full.names = T, all.files = T)
env126 = stack(bc126) 
env126 = check.env(env126, verbose = T)

bc245 = list.files("E:/Laura/TESIS/RSD/CAPAS/245", pattern = ".tif", full.names = T, all.files = T)
env245 = stack(bc245) 
env245 = check.env(env245, verbose = T)

bc370 = list.files("E:/Laura/TESIS/RSD/CAPAS/370", pattern = ".tif", full.names = T, all.files = T)
env370 = stack(bc370) 
env370 = check.env(env370, verbose = T)

bc585 = list.files("E:/Laura/TESIS/RSD/CAPAS/585", pattern = ".tif", full.names = T, all.files = T)
env585 = stack(bc585) 
env585 = check.env(env585, verbose = T)

bc1261 = list.files("E:/Laura/TESIS/RSD/CAPAS/1261", pattern = ".tif", full.names = T, all.files = T)
env1261 = stack(bc1261) 
env1261 = check.env(env1261, verbose = T)

bc2451 = list.files("E:/Laura/TESIS/RSD/CAPAS/2451", pattern = ".tif", full.names = T, all.files = T)
env2451 = stack(bc2451) 
env2451 = check.env(env2451, verbose = T)

bc3701 = list.files("E:/Laura/TESIS/RSD/CAPAS/3701", pattern = ".tif", full.names = T, all.files = T)
env3701 = stack(bc3701) 
env3701 = check.env(env3701, verbose = T)

bc5851 = list.files("E:/Laura/TESIS/RSD/CAPAS/5851", pattern = ".tif", full.names = T, all.files = T)
env5851 = stack(bc5851) 
env5851 = check.env(env5851, verbose = T)
```

Model prediction 
```{r}
ACTUAL = predict(
  modelRSD,
  env,
  type = 'logistic',
  file = "actual",
  format = "GTiff",
  overwrite = T
)
ACTUAL1 = predict(
  modelRSD,
  env2,
  type = 'logistic',
  file = "Actual1",
  format = "GTiff",
  overwrite = T
)
SSP126 = predict(
  modelRSD,
  env126,
  type = 'logistic',
  file = "ssp126",
  format = "GTiff",
  overwrite = T
)
SSP245 = predict(
  modelRSD,
  env245,
  type = 'logistic',
  file = "ssp245",
  format = "GTiff",
  overwrite = T
)
SSP370 = predict(
  modelRSD,
  env370,
  type = 'logistic',
  file = "ssp370",
  format = "GTiff",
  overwrite = T
)
SSP585 = predict(
  modelRSD,
  env585,
  type = 'logistic',
  file = "ssp585",
  format = "GTiff",
  overwrite = T
)
SSP1261 = predict(
  modelRSD,
  env1261,
  type = 'logistic',
  file = "ssp1261",
  format = "GTiff",
  overwrite = T
)
SSP2451 = predict(
  modelRSD,
  env2451,
  type = 'logistic',
  file = "ssp2451",
  format = "GTiff",
  overwrite = T
)
SSP3701 = predict(
  modelRSD,
  env3701,
  type = 'logistic',
  file = "ssp3701",
  format = "GTiff",
  overwrite = T
)
SSP5851 = predict(
  modelRSD,
  env5851,
  type = 'logistic',
  file = "ssp5851",
  format = "GTiff",
  overwrite = T
)
```

Plot prediction
```{r}
plotPred(ACTUAL, lt = "Actual", hr = T)
plotPred(ACTUAL1, lt = "Actual", hr = T)
plotPred(SSP126, lt = "SSP126", hr = T)
plotPred(SSP245, lt = "SSP245", hr = T)
plotPred(SSP370, lt = "SSP370", hr = T)
plotPred(SSP585, lt = "SSP585", hr = T)
plotPred(SSP1261, lt = "SSP126", hr = T)
plotPred(SSP2451, lt = "SSP245", hr = T)
plotPred(SSP3701, lt = "SSP370", hr = T)
plotPred(SSP5851, lt = "SSP585", hr = T)
```

Environmental response 
```{r}
plotResponse(
  modelRSD,
  "A1",
  type = NULL,
  only_presence = TRUE,
  marginal = F,
  rug = F,
  color = "red"
)
plotResponse(
  modelRSD,
  "A2",
  type = NULL,
  only_presence = TRUE,
  marginal = F,
  rug = F,
  color = "red"
)
plotResponse(
  modelRSD,
  "A3",
  type = NULL,
  only_presence = TRUE,
  marginal = F,
  rug = F,
  color = "red"
)
plotResponse(
  modelRSD,
  "A5",
  type = NULL,
  only_presence = TRUE,
  marginal = F,
  rug = F,
  color = "red"
)
```

ROC 
```{r}
plotROC(modelRSD, test = test)
```

```{r}
jk <- doJk(
  modelRSD,
  metric = "auc",
  test = test)

plotJk(
  jk,
  type = "train",
  ref = auc(modelRSD))

plotJk(
  jk,
  type = "test",
  ref = auc(modelRSD, test = test))
```

